---
title: "code_backplan"
author: "Yiling Yang"
date: "11/30/2019"
output: html_document
---

## suicide rate accross the age group
```{r}
maindata %>% 
ggplot(aes(x = suicideRate, fill = age, alpha = 0.5)) + 
geom_histogram() + 
facet_wrap(~age) +
labs(title = "Distribution of suicide rates across age groups", x = "Suicide rates")
```

comment:Lots of countries having low suicide rates for the youngest age group; some countries have high suicide rates when it comes to the older age groups

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# map
country <- maindata %>%
  group_by(country) %>% 
  summarise(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) * 100000)

countrydata <- joinCountryData2Map(country, joinCode = "NAME", nameJoinColumn = "country")

#par(mar=c(0, 0, 0, 0)) # margins

mapCountryData(countrydata, 
               nameColumnToPlot="suicide_per_100k", 
               mapTitle="", 
               colourPalette = "heat", 
               oceanCol="skyblue", 
               missingCountryCol="grey", 
               catMethod = "pretty")
```


```{r}
# bar for total
maindata%>% 
  group_by(country) %>% 
  summarize(count = n(), medianSuicides = median(suicides_no))  %>% 
  mutate(country = fct_reorder(country, medianSuicides)) %>% 
  top_n(20) %>% 
  ungroup() %>% 
  ggplot(aes(country, medianSuicides)) + 
  geom_col(aes(fill = country)) +
  coord_flip() +
  labs(title = "Median number of suicides per country", x = "Country", y = "Median suicides")
```


# Model

## model by best subset selection

```{r}
datatry= maindata_try %>% select(sex, age_group, suicides_no, log_suicide, gdp_for_year, gdp_per_capita, population)

full_reg = regsubsets(log_suicide~., data= datatry)
full_sum = summary(full_reg)
```

```{r}
rsq <- as.data.frame(full_sum$adjr2)
names(rsq) <- "R2"
rsq %>% 
        ggvis(x=~ c(1:nrow(rsq)), y=~R2 ) %>%
        layer_points(fill = ~ R2 ) %>%
        add_axis("y", title = "R2") %>% 
        add_axis("x", title = "Number of variables")

plot(full_reg,scale="adjr2")
```
By using $R^2$ as the criteria, BSS chose the model above.

```{r}
model_try = lm(log_suicide~sex+age_group+ suicides_no+ gdp_per_capita+ population, data = datatry)
summary(model_try)
plot(model_try)
```

after selection r squared increased a little but the residual plot was worse than before. We can choose this model because the model only keeps one of the two indicators for GDP.

### prediction for US

```{r}
us=maindata_try %>% filter(country=="United States")
hist(us$log_suicide)
modelus=lm(log_suicide~gdp_for_year+sex+age_group+gdp_per_capita, data = us)

summary(modelus)
plot(modelus)

# prediction for 25-34 male suicide rate in US
#sqrt(log(2.14*(10^13))*(-0.059)+1*0.07176+0.01109+65112*0.000001792+1.648)


```
The model has high R^2 but the residual plot was not good.


```{r}
country_mean_gdp <- maindata %>%
  group_by(country) %>%
  summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) * 100000, 
            gdp_per_capita = mean(gdp_per_capita))

ggplot(country_mean_gdp, aes(x = gdp_per_capita, y = suicide_per_100k)) + 
  geom_point() + 
  scale_x_continuous(labels=scales::dollar_format(prefix="$"), breaks = seq(0, 70000, 10000)) + 
  labs(title = "Correlation between GDP (per capita) and Suicides per 100k", 
       subtitle = "Plot containing every country",
       x = "GDP (per capita)", 
       y = "Suicides per 100k") 

model_gdp <- lm(suicide_per_100k ~ gdp_per_capita, data = country_mean_gdp)

gdp_suicide_no_outliers <- model_gdp %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  filter(.cooksd < 4/nrow(.)) %>% # removes 5/93 countries
  inner_join(country_mean_gdp, by = c("suicide_per_100k", "gdp_per_capita")) 

model_gdp_ad <- lm(suicide_per_100k ~ gdp_per_capita, data = gdp_suicide_no_outliers)

summary(model_gdp_ad)


ggplot(gdp_suicide_no_outliers, aes(x = gdp_per_capita, y = suicide_per_100k)) + 
  geom_point() + 
  geom_smooth(method = "lm", aes(group = 1)) + 
  scale_x_continuous(labels=scales::dollar_format(prefix="$"), breaks = seq(0, 70000, 10000)) + 
  labs(title = "Correlation between GDP per capita and Suicides per 100k", 
       subtitle = "Plot with high CooksD countries removed (5/93 total)",
       x = "GDP per capita", 
       y = "Suicides per 100k", 
       col = "Continent") + 
  theme(legend.position = "none")
```

There is weak positive relationship between GDP per capita and suicides per 100k after we removed outliers.
